## Effect of Treatment Over Time

The `PrePost` dataset contains measurements taken before and after a treatment training. The same participants were measured at both time points, allowing us to examine whether the measured outcome changed after the treatment. Because the before and after measurements come from the same participants, the observations are paired.

### Import data

```{r}

# 1.Import the data
  
prepost <- read_csv("data/PrePost copy.csv")

  # View the first 6 rows to make sure it was imported correctly 
  head(prepost)

```

### Check the structure of the data

8 measurements before rehabilitation and 8 after:

```         
Before: 125 130 132 135 136 138 140 145
After:  127 132 133 136 139 141 145 148
```

```{r}

# Display the variable names
names(prepost)

# Check the structure and data types
str(prepost)

# Obtain a first summary of the values
summary(prepost)
```

### Variables of the dataset

| **Variable** | **Description** | **Unit** | **Possible Values** | **Expected Change** |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| `perf` | Performance measurement taken before and after cardiac rehabilitation | Not specified | Numerical values; observed in this dataset from 125 to 148 | Expected to increase if a higher performance score represents better physical capacity |
| `time` | Indicates whether the performance measurement was taken before or after rehabilitation | None | `Before`, `After` | Not applicable; this is the measurement occasion |

### Data analysis

Q: Does the rehabilitation treatment have an effect on measured performance from before to after treatment?

**a) What comparison are we making?**

- performance before rehabilitation vs performance after rehabilitation

  - Before: 125 130 132 135 136 138 140 145
  - After: 127 132 133 136 139 141 145 148

- If the same 8 participants are being compared before and after then this is a paired comparison

**b) Parametric or non-parametric test?**

- For a paired test, what matters is the distribution of the differences:
  - $difference = After - Before$
  - The differences are \[2 2 1 1 3 3 5 3\]
- Check that the differences are normally distributed using a `Shapiro-Wilk normality test`.
  - If $p > 0.05$ = not statistically significant and there is not enough evidence to say the differences depart from normality (we do not reject normality) $\rightarrow$ **`parametric test`** $\rightarrow$ `paired t-test` (Kim, 2015).
  - If $p < 0.05$ = statistically significant and there is enough evidence to say the differences are not normally distributed (we reject normality) $\rightarrow$ **`non-parametric test`** $\rightarrow$ **`Wilcoxon signed-rank test`** (Whitley & Ball, 2002).
  - For this dataset $p = 0.283 > 0.05$ therefore we do not reject normality of the differences and a parametric paired t-test can be used for statistical analysis since the data are paired (Kim, 2015).
- The `paired t-test` gives $p = 0.001 < 0.05$ $\rightarrow$ there is a statistically significant difference between performance before and after rehabilitation.

```{r}
# 1. Calculate the differences between before and after in the dataset

# a) Separate the before and after performance values
before <- prepost %>% filter(time == "Before") %>% pull(perf) 
# only keep rows where time == "Before" and extract the perf column 

after <- prepost %>% filter(time == "After") %>% pull(perf)
# only keep rows where time == "After" and extract the perf column 

# b) Calculate the difference for each participant
difference <- after - before

difference

```

```{r}
# 2. Check for normality of data using Shapiro-Wilk normality test

shapiro.test(difference)

```

```{r}
# 3. Paired t-test comparing performance before and after rehabilitation
t.test(after, before, paired = TRUE)
```

**c) Which graph illustrates the effect of rehabilitation treatment?**

Since the same participants were measured before and after the rehabilitation treatment, a paired line plot was used. This type of graph connects each participant's before and after measurements, making it easy to see the direction and size of the change for each individual rather than only comparing the overall group averages.

```{r}

# 1. Reorganize data structure so each participant has a id (number 1-8)

# a). Add a new column to dataset: participant IDs to label each participant 
prepost_id <- prepost %>% mutate(id = rep(1:8, 2), time = factor(time, levels = c("Before", "After")))  # give participants a # 1-8 and repeat twice. Before comes first and after comes second.
 

# 2. Create the paired line plot
ggplot(prepost_id, aes(x = time, y = perf, group = id, color = factor(id))) + 
  #color = factor(id) → converts numbers to categorical data to label each person with a colour
  # group = id → connect the correct person's points
  
  geom_line() +
  geom_point() +
  labs(title = "Performance Before and After Rehabilitation", 
       x = "Time", y = "Performance", color = "Participant") + # create legend labeled participant
  theme_minimal()


```

**d) Rehabilitation treatment interpretation**

Mean performance increased after rehabilitation compared with before. The slightly larger standard deviation after rehabilitation indicates that participant performance values were slightly more spread out around the mean.

- **Before rehabilitation:** `135.1 ± 6.2` = mean ± standard deviation

- **After rehabilitation:** `137.6 ± 7.0` = mean ± standard deviation

  - Smaller SD = participants values are more similar to each other

  - Larger SD = participant values are more spread out.

```{r}

# 1. Calculate the mean & SD before rehabilitation

mean(before)
sd(before)

# 2. Calculate the mean & SD after rehabilitation

mean(after)
sd(after)
```
