## Effect of Treatment Over Time

This dataset investigates whether the non-use of the proximal upper limb (PANU) is related to the non-use of two of its component joints: the shoulder (SANU) and the elbow (EENU). Non-use refers to the extent to which a patient avoids using a joint or limb segment during daily activities, typically measured in clinical settings after neurological injury (e.g., stroke). Each variable is expressed as a percentage, where higher values indicate greater non-use. The question is whether limiting the use of a single joint (shoulder or elbow) is associated with limiting the use of the whole arm (PANU).

### a). Import data

```{r}

# 1.Import the data
  
  prepost <- read_csv("../data/PrePost copy.csv")
  
  # View the first 6 rows to make sure it was imported correctly 
  head(prepost)

```

### b). Check the structure of the data

8 measurements before rehabilitation and 8 after:

```         
Before: 125 130 132 135 136 138 140 145
After:  127 132 133 136 139 141 145 148
```

```{r}

# Display the variable names
names(prepost)

# Check the structure and data types
str(prepost)

# Obtain a first summary of the values
summary(prepost)
```

### c). Variables of the dataset

| **Variable** | **Description** | **Unit** | **Possible Values** | **Expected Change** |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| `perf` | Performance measurement taken before and after cardiac rehabilitation | Not specified | Numerical values; observed in this dataset from 125 to 148 | Expected to increase if a higher performance score represents better physical capacity |
| `time` | Indicates whether the performance measurement was taken before or after rehabilitation | None | `Before`, `After` | Not applicable; this is the measurement occasion |

### d). Data analysis

Q: Does the rehabilitation treatment have an effect on measured performance from before to after treatment?

**a) What comparison are we making?**

- performance before rehabilitation vs performance after rehabilitation

  - Before: 125 130 132 135 136 138 140 145
  - After: 127 132 133 136 139 141 145 148

- If the same 8 participants are being compared before and after then this is a paired comparison

**b) Parametric or non-parametric test?**

- For a paired test, what matters is the distribution of the differences:
  - $difference = After - Before$
  - The differences are \[2 2 1 1 3 3 5 3\]
- Check that the differences are normally distributed using a `Shapiro-Wilk normality test`.
  - If $p > 0.05$ = not statistically significant and there is not enough evidence to say the differences depart from normality (we do not reject normality) $\rightarrow$ **`parametric test`** $\rightarrow$ `paired t-test`
  - If $p < 0.05$ = statistically signifcant and there is enough evidence to say the differences are not normally distributed (we reject normality) $\rightarrow$ **`non-parametric test`** $\rightarrow$ **`Wilcoxon signed-rank test`**.
  - For this dataset $p = 0.283 > 0.05$ therefore the data is normally distributed and a parametric paired t-test can be used for statistical analysis since the data are paired.
- The `paired t-test` gives $p = 0.001 < 0.05$ $\rightarrow$ there is a statistically significant difference between performance before and after rehabilitation.

```{r}
# 1. Calculate the differences between before and after in the dataset

# a) Separate the before and after performance values
before <- prepost %>% filter(time == "Before") %>% pull(perf) 
# only keep rows where time == "Before" and extract the perf column 

after <- prepost %>% filter(time == "After") %>% pull(perf)
# only keep rows where time == "After" and extract the perf column 

# b) Calculate the difference for each participant
difference <- after - before

difference

```

```{r}
# 2. Check for normality of data using Shapiro-Wilk normality test

shapiro.test(difference)

```

```{r}
# 3. Paired t-test comparing performance before and after rehabilitation
t.test(after, before, paired = TRUE)
```

**c) Which graph illustrates the effect of rehabilitation treatment?**

Since this is paired data, use a paired line plot to best show the effect of rehabilitation treatment on each participant.

```{r}

# 1. Reorganize data structure so each participant has a id (number 1-8)

# a). Add a new column to dataset: participant IDs to label each participant 
prepost_id <- prepost %>% mutate(id = rep(1:8, 2), time = factor(time, levels = c("Before", "After")))  # give participants a # 1-8 and repeat twice. Before comes first and after comes second.
 

# 2. Create the paired line plot
ggplot(prepost_id, aes(x = time, y = perf, group = id, color = factor(id))) + 
  #color = factor(id) → converts numbers to categorical data to label each person with a colour
  # group = id → connect the correct person's points
  
  geom_line() +
  geom_point() +
  labs(title = "Performance Before and After Rehabilitation", 
       x = "Time", y = "Performance", color = "Participant") + # create legend labeled participant
  theme_minimal()


```

**d) Rehabilitation treatment interpretation**

Mean performance increased after rehabilitation compared with before. The slightly larger standard deviation after rehabilitation indicates that participant performance values were slightly more spread out around the mean.

- **Before rehabilitation:** `135.1 ± 6.2` = mean ± standard deviation

- **After rehabilitation:** `137.6 ± 7.0` = mean ± standard deviation

  - Smaller SD = participants values are more similar to each other

  - Larger SD = participant values are more spread out.

```{r}

# 1. Calculate the mean & SD before rehabilitation

mean(before)
sd(before)

# 2. Calculate the mean & SD after rehabilitation

mean(after)
sd(after)
```