

## Effect of Treatment Over Time

This dataset investigates whether the non-use of the proximal upper limb (PANU) is related to the non-use of two of its component joints: the shoulder (SANU) and the elbow (EENU). Non-use refers to the extent to which a patient avoids using a joint or limb segment during daily activities, typically measured in clinical settings after neurological injury (e.g., stroke). Each variable is expressed as a percentage, where higher values indicate greater non-use. The question is whether limiting the use of a single joint (shoulder or elbow) is associated with limiting the use of the whole arm (PANU).

### The Data


```{r load-data}
nonuse <- read.csv("../data/NonUse copy.csv")
summary(nonuse)
```

#### Measured Variables

| variable | description | unit | range |
|----------|-------------|------|-------|
| EENU | non use of the elbow extension | percentage | [0%..100%] |
| SANU | non use of shoulder flexion/antepulsion | percentage | [0%..100%] |
| PANU | non use of proximal arm joint (shoulder and elbow) | percentage | [0%..100%] |

### The Analysis

#### Does the Treatment Have an Effect?

Yes: performance is higher after the treatment (see the tests below).

#### What Comparison Are We Making?

We compare the performance of the same patients before and after the treatment, so this is a **paired** comparison. The data contain no subject identifier, so we assume that the row order keeps the Before/After pairing (8 patients, 2 measurements each).


#### Parametric or Non-Parametric?

The paired t-test assumes that the differences are normally distributed. We check this with a Shapiro-Wilk test.

```{r prepost-test}
prepost <- prepost %>%
  mutate(subject = rep(1:8, times = 2),
         time = factor(time, levels = c("Before", "After")))

before <- prepost$perf[prepost$time == "Before"]
after  <- prepost$perf[prepost$time == "After"]
diff   <- after - before

shapiro.test(diff)
t.test(before, after, paired = TRUE)
wilcox.test(before, after, paired = TRUE)
```


Shapiro-Wilk gives p = 0.283 (> 0.05), so the differences are compatible with a normal distribution. The **paired t-test** (parametric) is therefore appropriate. The Wilcoxon test (non-parametric) gives the same conclusion (p = 0.014).

#### Which Graph Illustrates the Effect?

#### Thesis-Style Results Sentence