## Testing Some Stereotypes

### Are Snorers Fatter?

#### a) Import data

```{r}

# 1.Import the data

library(readr)

snore <- read_delim("../data/snore copy.csv") # delim seperates the columns correctly 

# View the first 6 rows to make sure it was imported correctly 
head(snore)
```

#### b) Comparison Variable

The BMI of snorers and non-snores can be compared to gives us a measure of body mass relative to height, so it is more appropriate than simply asking whether snorers weigh more.

$BMI_{snorers}$ vs $BMI_{non-snoreres}$

$BMI = \frac{weight(kg)}{height (m)^2}$

```{r}

# 1. Create BMI variable: create a new column BMI with the height m 
snore <- snore %>% mutate(BMI = weight / (height / 100)^2) # mutate adds a new column BMI 

head(snore)

```

#### c) Descriptive Analysis

```{r}
# 2. # Visualize BMI distribution: boxplot before deciding the type of statistical test

ggplot(snore, aes(x = snore, y = BMI)) +
  geom_boxplot() +
  labs(title = "BMI of Snorers and Non-Snorers", x = "Snoring", y = "BMI (kg/m²)") +
  theme_minimal()
```

```{r}

# 3. Descriptive statistics table
snore %>% group_by(snore) %>%
  summarise(mean_BMI = mean(BMI), median_BMI = median(BMI), sd_BMI = sd(BMI), n = n())
```

Both groups are extremely similar and the boxplot also shows a very similar distribution and spread in both groups. Therefore descriptively, there is no obvious evidence that snorers have a higher BMI.


