## Testing Some Stereotypes

### Are Snorers Fatter?

#### a) Import data

```{r}

# 1.Import the data

library(tidyverse)

snore <- read_delim("../data/snore copy.csv") # delim seperates the columns correctly 

# View the first 6 rows to make sure it was imported correctly 
head(snore)
```

#### b) Comparison Variable

The BMI of snorers and non-snores can be compared to gives us a measure of body mass relative to height, so it is more appropriate than simply asking whether snorers weigh more.

$BMI_{snorers}$ vs $BMI_{non-snoreres}$

$BMI = \frac{weight(kg)}{height (m)^2}$

```{r}

# 1. Create BMI variable: create a new column BMI with the height m 
snore <- snore %>% mutate(BMI = weight / (height / 100)^2) # mutate adds a new column BMI 

head(snore)

```

#### c) Descriptive Analysis

```{r}
# 2. # Visualize BMI distribution: boxplot before deciding the type of statistical test

ggplot(snore, aes(x = snore, y = BMI)) +
  geom_boxplot() +
  labs(title = "BMI of Snorers and Non-Snorers", x = "Snoring", y = "BMI (kg/m²)") +
  theme_minimal()
```

```{r}

# 3. Descriptive statistics table
snore %>% group_by(snore) %>%
  summarise(mean_BMI = mean(BMI), median_BMI = median(BMI), sd_BMI = sd(BMI), n = n())
```

Both groups are extremely similar and the boxplot also shows a very similar distribution and spread in both groups. Therefore descriptively, there is no obvious evidence that snorers have a higher BMI.

- Non-snorers (`N`): mean BMI = 27.27, median = 28.07, SD = 2.81, (n=65)

- Snorers (`O`): mean BMI = 27.11, median = 28.08, SD = 2.83, (n=35)

These are **unpaired/independent groups**, because the snorers and non-snorers are different people.

Before deciding the type of statistical test: check to see if BMI is normally distributed in each group using the Shapiro-Wilk test.

#### d) Statistical Test Selection

```{r}
# 4. Check the normality of BMI in both groups using Shapiro-Wilk test.

# a) Check normality of BMI for non-snorers
shapiro.test(snore$BMI[snore$snore == "N"]) # take the BMI column only for the rows where snore = no 

# b) Check normality of BMI for snorers
shapiro.test(snore$BMI[snore$snore == "O"]) # take the BMI column only for the rows where snore = yes

```

BMI is not normally distributed in either group. Both are much smaller than (0.05), so we reject the assumption of normality for both groups:

$BMI_{snoreres}$ $\rightarrow$ $p = 2.323e-08 < 0.05$

$BMI_{non-snoreres}$ $\rightarrow$ $p = 2.323e9.81-07 < 0.05$

Since the two groups are independent and the data are non-normal, use the `Wilcoxon rank-sum test` (also called the `Mann–Whitney U test`). This tests whether the BMI distributions differ between snorers and non-snorers.

#### e) Statistical Comparison

```{r}

# 5. Run the Wilcoxon rank-sum test on BMI 

wilcox.test(BMI ~ snore, data = snore)
```

$p = 0.951 > 0.05$ $\rightarrow$ there is no statistically significant difference in BMI between snorers and non-snorers. The BMI distributions of snorers and non-snorers are not meaningfully different in this sample.

#### f) Conclusion

BMI was similar between snorers and non-snorers, with median values of 28.08 and 28.07 kg/m², respectively. The difference was not statistically significant (Wilcoxon rank-sum test, W = 1128.5, p = 0.951), providing no evidence to support the stereotype that snorers are fatter.
