

### Do Snorers Drink or Smoke More?

 **Do snorers drink more?**

Alcohol consumption of snorers and non-snorers can be compared to determine whether snorers have higher alcohol consumption. Analyze the distribution of alcohol consumption before statistical analyses

$Alcohol_{snorers}$ vs $Alcohol_{non-snorers}$

```{r}
# 1. Make a boxplot to compare alcohol consumption between snorers and non-snorers
ggplot(snore, aes(x = snore, y = alcool)) +
  geom_boxplot() +
  labs(title = "Alcohol Consumption of Snorers and Non-Snorers", x = "Snoring", y = "Alcohol") +
  theme_minimal()
```

```{r}
# 2. Create a table to look at the descriptive statistics

snore %>%
  group_by(snore) %>%
  summarise(mean_alcool = mean(alcool), median_alcool = median(alcool), sd_alcool = sd(alcool),   n = n())
```

Snorers appear to consume more alcohol than non-snorers. Snorers had a higher mean alcohol consumption (4.03 vs 2.37) and a higher median (4 vs 0). The boxplot also shows generally higher alcohol values among snorers.

- Non-snorers (`N`): mean alcohol = 2.37, median = 0, SD = 3.18

- Snorers (`O`): mean alcohol = 4.03, median = 4, SD = 3.47

```{r}
# 3. Check the normality of alcohol consumption in both groups using the Shapiro-Wilk test

# a) Check normality for non-snorers
shapiro.test(snore$alcool[snore$snore == "N"])

# b) Check normality for snorers
shapiro.test(snore$alcool[snore$snore == "O"])
```

Alcohol consumption was not normally distributed in either snorers or non-snorers (Shapiro–Wilk tests, both p \< 0.05). Therefore, a non-parametric test was used to compare the two unpaired/independent groups.

$Alcohol_{snoreres}$ $\rightarrow$ $p = 6.818e-09 < 0.05$

$Alcohol_{non-snoreres}$ $\rightarrow$ $p = 0.005032 < 0.05$

```{r}
# 4. Run the Wilcoxon rank-sum test on alcohol consumption

wilcox.test(alcool ~ snore, data = snore)
```


This result is statistically significant, meaning there is a difference between the two groups.

$p = 0.008594 < 0.05$

Snorers had higher alcohol consumption than non-snorers, with median values of 4 and 0, respectively. This difference was statistically significant (Wilcoxon rank-sum test, W = 788, p = 0.009), supporting the stereotype that snorers drink more.

