### Are Men Fatter?

The `snore` dataset is also used to compare BMI between men and women. BMI is calculated from each participant's height and weight, and the two sex groups are treated as independent groups. The distribution of BMI is first examined before choosing the appropriate statistical test.

A box plot is first used to compare the BMI distributions of men and women before performing the statistical analysis.

$BMI_{men}$ vs $BMI_{women}$

```{r}
# 1. Make a boxplot to compare BMI between men and women

ggplot(snore, aes(x = sex, y = BMI)) +
  geom_boxplot() +
  labs(title = "BMI of Men and Women", x = "Sex", y = "BMI (kg/m²)") +
  theme_minimal()
```

```{r}
# 2. Create a table to look at the descriptive statistics

snore %>%
  group_by(sex) %>%
  summarise(mean_BMI = mean(BMI), median_BMI = median(BMI), sd_BMI = sd(BMI), n = n())
```

Descriptive results suggest that men have a slightly higher BMI than women, but the difference looks small. The median BMI was 28.08 kg/m² for men and 27.73 kg/m² for women, while the mean BMI was 27.27 and 27.04 kg/m², respectively.

- Women (`F`): mean BMI = **27.04**, median = **27.73**, SD = **2.72**, (n=25)

- Men (`H`): mean BMI = **27.27**, median = **28.08**, SD = **2.85**, (n=75)

Check the normality of the data using the Shapiro-Wilk test (Ghasemi & Zahediasl, 2012)

```{r}

# 3. Check normality of BMI in both groups using the Shapiro-Wilk test

# a) Women
shapiro.test(snore$BMI[snore$sex == "F"])

# b) Men
shapiro.test(snore$BMI[snore$sex == "H"])
```

BMI was not normally distributed in either women or men (Shapiro–Wilk tests, both p \< 0.05). Therefore, a non-parametric test was used to compare BMI between the two independent groups.

- Women: $p = 1.154e-05 < 0.05$

- Men: $p = 3.456e-09 < 0.05$

```{r}

# 4. Run the Wilcoxon rank-sum test on BMI

wilcox.test(BMI ~ sex, data = snore)
```

No statistically significant difference in BMI between men and women.

- $p = 0.488 > 0.05$

Men had a slightly higher median BMI than women (28.08 vs 27.73 kg/m²), but the difference was not statistically significant (Wilcoxon rank-sum test, W = 850, p = 0.488). Therefore, the data do not support the stereotype that men are fatter.
