### Do Women Smoke Less?

The `snore` dataset is used to compare smoking status between women and men. Since both sex and smoking status are categorical variables, the analysis focuses on comparing the proportion of smokers in each group.

#### Comparison Variables

Both `sex` and `smoking` are categorical variables. Therefore, the proportion of smokers is compared between women and men.

$Smokers_{men}$ vs $Smokers_{women}$

#### Descriptive Analysis

```{r}

# 1. Make a proportional bar plot to compare smoking status between women and men

ggplot(snore, aes(x = sex, fill = tabac)) +
  geom_bar(position = "fill") +
  labs(title = "Smoking Status of Women and Men", x = "Sex", y = "Proportion", 
  fill = "Smoking") +
  theme_minimal()
```

```{r}
# 2. Create a contingency table for sex and smoking status

sex_smoke_table <- table(snore$sex, snore$tabac)

# a) Calculate proportions within each sex

prop.table(sex_smoke_table, margin = 1)

# b) Convert the proportions to percentages

prop.table(sex_smoke_table, margin = 1) * 100
```

A lower proportion of women were smokers compared with men (40% vs 72%). Descriptively, this supports the stereotype that women smoke less.

- **Women (`F`)**: 40% smokers, 60% non-smokers

- **Men (`H`)**: 72% smokers, 28% non-smokers

#### Statistical Test Selection

Both `sex` and `tabac` are categorical variables, and the observations are independent because each participant belongs to only one sex and one smoking-status category.

A `Chi-square test` of independence can therefore be used to test whether smoking status is associated with sex.

For the Chi-square test, the expected cell counts must be reasonably large. If the expected counts are all around 5 or more, the Chi-square test is appropriate (Illowsky, 2023).

```{r}
# 3. Test whether smoking status is associated with sex

sex_smoke_chi <- chisq.test(sex_smoke_table)

sex_smoke_chi

# Check expected counts for the Chi-square test
sex_smoke_chi$expected

```

The expected counts were sufficiently large, so the assumptions of the Chi-square test were considered satisfied.

#### Conclusion

There is significant association between sex and smoking status.

- $p=0.008<0.05$

A lower proportion of women were smokers compared with men (40% vs 72%). This difference was statistically significant ($\chi^2(1)=7.00, p=0.008$), supporting the stereotype that women smoke less in this sample.
