

### Do Women Smoke Less?

#### a) Comparison Variables

Both gender and `smoking` are categorical variables therefore the proportion of smokers among women and men must be compared.

$Smokers_{men}$ vs $Smokers_{women}$

#### b) Descriptive Analysis

```{r}

# 1. Make a proportional bar plot to compare smoking status between women and men

ggplot(snore, aes(x = sex, fill = tabac)) +
  geom_bar(position = "fill") +
  labs(title = "Smoking Status of Women and Men", x = "Sex", y = "Proportion", 
  fill = "Smoking") +
  theme_minimal()
```

```{r}
# 2. Create a contingency table for sex and smoking status

sex_smoke_table <- table(snore$sex, snore$tabac)

# a) Calculate proportions within each sex

prop.table(sex_smoke_table, margin = 1)

# b) Convert the proportions to percentages

prop.table(sex_smoke_table, margin = 1) * 100
```

A lower proportion of women were smokers compared with men (40% vs 72%). Descriptively, this supports the stereotype that women smoke less.

- **Women (`F`)**: 40% smokers, 60% non-smokers

- **Men (`H`)**: 72% smokers, 28% non-smokers

#### c) Statistical Test Selection

Both `sex` and `tabac` are categorical variables, and the observations are independent because each participant belongs to only one sex and one smoking-status category.

A Chi-square test of independence can therefore be used to test whether smoking status is associated with sex.

For the Chi-square test, the expected cell counts must be reasonably large. If the expected counts are all around 5 or more, the Chi-square test is appropriate.

```{r}
# 3. Test whether smoking status is associated with sex

sex_smoke_chi <- chisq.test(sex_smoke_table)

sex_smoke_chi

# Check expected counts for the Chi-square test
sex_smoke_chi$expected

```
The expected counts were sufficiently large, so the assumptions of the Chi-square test were considered satisfied.

