
### Are There Any Correlations Between Variables?

8.From a more general perspective: are there any correlations between variables?**

We are no longer comparing two groups. We want to see whether the numerical variables tend to change together.

- **`Numerical variables`:** age, weight, height, alcohol

- **`Categorical variables`:** snore, gender, smoking

Categorical data should not be put in a standard correlation matrix since they are not numerical quantities. Based on the previous analysis above, some numerical variables were not normally distributed, therefore `Spearman correlation` can be used to see whether two variables generally increase or decrease together. Spearman is used instead of Pearson because Spearman does not require normality.

1.  A dataset must be created for only the numerical variables.
2.  The Spearman correlation matrix is calculated.
3.  The interesting correlations are then tested to see if they are statistically significant or not.

```{r}
# 1. Select the numerical variables for the correlation analysis

snore_numeric <- snore %>%
  select(age, weight, height, alcool)

head(snore_numeric)
```

```{r}
# 2. Calculate Spearman correlations between all numerical variables

correlation_matrix <- cor(snore_numeric, method = "spearman", use = "complete.obs")

correlation_matrix
```

The main relationship is between weight and height. It is a very strong positive correlation. Taller participants generally tend to weigh more. All other correlations were weak or close to zero.

- $r_s = 0.930$

- `1.0` = variables are perfectly correlated with each other

Now the correlation between wweight and height can be tested to see whether it is test statistically significant or not. `Spearman correlation` is used to test the significance. Spearman correlation was used because some numerical variables were not normally distributed as mentioned previously, making a non-parametric correlation more appropriate.

```{r}
# 3. Test the correlation between height and weight

cor.test(snore$height, snore$weight, method = "spearman")
```

There is a very strong positive and statistically significant association between height and weight: as height increases, weight generally increases as well. $r_s = 0.930$

- $p < 2.2e-16$

```{r}
# 4. Visualize the relationship between height and weight

ggplot(snore, aes(x = height, y = weight)) +
  geom_point() +
  labs(title = "Relationship Between Height and Weight", x = "Height", y = "Weight") +
  theme_minimal()
```

Height and weight showed a very strong positive correlation ($r_s = 0.930, p = 2.2e-16$), indicating that taller participants generally had greater body weight. The other numerical variables showed only weak or negligible correlations.

